Classless Java

### 2016-03-23 OOPSLA16/sections/Implementation.tex
Notation ok?

\item 
Lombok offers only limited/experimental support for separate compilation, that is
to access information of code defined in different files.
In the same spirit, we have a mature \mixin annotation, that
does not support separate compilation yet: it requires that all
  related interfaces have to appear in a single Java file.
We also offer a less polished annotation, 
that trying to reuse the (yet experimental) Lombock annotation \lstinline{@Delegate}, reused the name and modified its behavior.
  Now files in the same package can be compiled and \lstinline{@Delegate} has the same behavior as \mixin. A limitation is
  that users need to put references to the super types of the annotated type together with the annotation, for instance,
  \lstinline{@Delegate(types = Point2D.class)}. Another limitation is the methods generated by \lstinline{@Delegate} will
  not be visible in the Outline window, but they can still be auto-completed in Eclipse.
\end{itemize}

>>>>

\item 
Lombok offers only limited/experimental support for separate compilation, that is
to access information of code defined in different files.
In the same spirit, we have a mature \mixin annotation, that
does not support separate compilation yet: it requires that all
  related interfaces have to appear in a single Java file.
Reusing the logic inside the yet experimental Lombock annotation \lstinline{@Delegate},
we also offer a less polished annotation supporting
separate compilation for files in the same package.
%  has the same behavior as \mixin. A limitation is
%  that users need to put references to the super types
% of the annotated type together with the annotation, for instance,
%  \lstinline{@Delegate(types = Point2D.class)}.
%   Another limitation is the methods generated by \lstinline{@Delegate} will
%  not be visible in the Outline window, but they can still be auto-completed in Eclipse.
\end{itemize}

### 2016-03-23 OOPSLA16/sections/Translation.tex
Notation ok?

This section gives an overview of what $\mixinAnn$ generates in
Classless Java, and what properties are guaranteed in the translation.

@@

Figure~\ref{figure:valid} and Figure~\ref{figure:ofmethod} present more details on the auxiliary
functions. \bruno{if the figures are mentioned here, what should
  follow is the explanation on what is one the figures. Please fix
  this Haoyuan and makes sure that everything that is on the figures
  is explained here. }
The formal definition is available in
Appendix~\ref{subsec:translation}.\bruno{what formal definition?}

The function $\ofMethod$ generates the static method $\QM{of}$, as an
object factory. It detects all the field methods of $\C_0$, take them
as arguments, and in the return statement is an anonymous class which
generates getters, fluent setters and with-methods in certain
cases. Some other features of $\mixinAnn$, including non-fluent
setters and the $\QM{with}$ method, however, are not formalized in
Classless Java. Appendix~\ref{subsec:otherfeatures} gives an informal
explanation of those methods.

\bruno{I think there is explanation missing in this subsection;
  probably text which is now in the appendix. Haoyuan, please take
  care of this.}

@@

To formally characterize the behavior of our annotation and the two levels of guarantees that we offer, we provide some notations and two theorems:
\begin{itemize}
\item We denote with $\C^{\II}$ the name of an interface.
\item An interface table
IT is OK if under such interface table, all interfaces are OK,
\marcoT{that is, well typed}.
\item Since interface tables are just represented as sequences of interfaces we write IT = $\II$ IT' to select a specific interface in a table.
\item IT contains an heir of $\C$ if there is an interface that extends it, or a \Q@new@ that instantiate it.
\end{itemize}

\bruno{Do we have good names for the theorems Marco? Cannot be called
  tunning now!}
\begin{thm}[@ObjOf]
If a given interface table $\II$ IT is OK
 where $\II$ has $\weakAnn$,
$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,
then the interface table $[\![\II]\!]$ IT is OK.
\end{thm}

>>>>

This section gives an overview of what $\mixinAnn$ generates,
 and what formal properties are guaranteed in the translation.

@@

Figure~\ref{figure:valid} and Figure~\ref{figure:ofmethod} present more details on the auxiliary
functions. 
The first two points of Figure~\ref{figure:valid}
defines function $\otherMethod$;
 This function generates unimplemented
 \Q@with-@ and fluent setters in the interface, where the 
return types has been refined.

To determine whether a method need to be generated,
we check if such \Q@with-@ or setter methods
require an implementation in $\C_0$, but not declared directly in $\C_0$.

The third point gives the definition
of $\valid$:
It is valid to annotate an interface if 
%To generate the \Q@of@ method, it is necessary for $\weakAnn$ to explicitly check if
all abstract methods (that is, all those
requiring an implementation) are valid.
That is, we can categorize them in a pattern that we
know how to implement (right column):
it is either a field getter (first point),
a with method (second point) or a setter (third point).

Note that we write $\QM{with#}\m$ to append $\m$ to
\QM{with}, following the camelCase rule. The first letter of $\m$
must be lower-case and is changed to upper-case upon appending. For
example \QM{with#foo}=\QM{withFoo}.  Special names $\specialName(\m)$
are \QM{with} and all identifiers of the form $\QM{with#}\m$.

Figure~\ref{figure:ofmethod},
defines the $\ofMethod$  function, that generates the static method $\QM{of}$ as an
object factory.
It detects all the field methods of $\C_0$ and use them to synthesize
its arguments.
The return statement instantiate an anonymous class which
generates the needed getters, fluent setters and with-methods.

The right column first point collect the getter methods,
the second and third point generate implementations for with methods
if needed; similarly, the fourth and fifth point generate fluent setters if needed.

Some other features of $\mixinAnn$, including non-fluent
setters and the $\QM{with}$ method are not formalized here.
Appendix~\ref{subsec:otherfeatures} gives an detailed but 
informal explanation of generation for those methods.

@@

To formally characterize the behavior of our annotation and the two levels of guarantees that we offer, we provide some notations and two theorems:
\begin{itemize}
\item We denote with $\C^{\II}$ and $m^\method$ the name of an interface and of a method.
\item An interface table
IT is OK if under such interface table, all interfaces are OK,
\marcoT{that is, well typed}.
\item Since interface tables are just represented as sequences of interfaces we write IT = $\II$ IT' to select a specific interface in a table.
\item IT contains an heir of $\C$ if there is an interface that extends it, or a \Q@new@ that instantiate it.
\end{itemize}

\begin{thm}[@ObjOf]
If a given interface table $\II$ IT is OK
 where $\II$ has $\weakAnn$,
$\valid(\C^{\II})$  and $\QM{of}\notin\dom(\C^{\II})$,
then the interface table $[\![\II]\!]$ IT is OK.
\end{thm}
